---
title: "Plots 3D data"
author: "Joao Marreiuros and David Nora"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---

# Goal of the script
This script reads and plots all the 3D edge reduction data

```{r}
dir_in <- "../raw_data"
dir_out <- "../plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.

---

# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(patchwork)
library(doBy)
library(ggrepel)
library(openxlsx)
library(flextable)
library(janitor)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.csv$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
info_in
```

# Load data db 
```{r}
imp_data <- read_csv("../raw_data/3dedge.csv")
str(imp_data)

```


# Summarize data

```{r}

# General overview

nminmaxmeanmedsd <- function(x){
	y <- x[!is.na(x)]
	n_test <- length(y)
	min_test <- min(y)
	max_test <- max(y)
	mean_test <- mean(y)
 	med_test <- median(y)
 	sd_test <- sd(y)
 	out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test)
 	names(out) <- c("n", "min", "max", "mean", "median", "sd")
 	return(out)
}

num.var <- 6:length(imp_data)

stats <- summaryBy(.~raw.material + cycle + ID, data=imp_data[c("raw.material","cycle", "ID", names(imp_data)[num.var])], FUN=nminmaxmeanmedsd)

write_csv(stats, "../derived_data/stats.csv")


# Summarizin number of parts per cycle, per raw material

imp_data$cycle <- factor(imp_data$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
imp_data$raw.material <- factor(imp_data$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))

parts <- imp_data %>%
  group_by(raw.material, cycle) %>%
  summarize(total = n()) %>%
  pivot_wider(names_from = "cycle",
              values_from = "total",
              values_fill = 0) %>%
  adorn_totals(where = c("row", "col"), fill = "") %>%
  rename("Raw material" = raw.material)

print(parts)

write_csv(parts, "../derived_data/parts.csv")



```

# Plots sumarized data

## Plot Max distance between samples

```{r}

# gets new order 
stats$cycle <- factor(stats$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
stats$raw.material <- factor(stats$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))



stats.plot <- ggplot(data = stats) +  
              geom_point(mapping = aes(x = raw.material, y = end.max, colour = raw.material)) +
              theme_classic() +
              labs(colour = "Raw material", x = "Raw material", y = "Maximum distance (mm)", title = "Max distance") +
              facet_wrap(~ cycle)
              
print(stats.plot)

ggsave("../plots/endmax.png")

```

## Plot distance intervals and number of counts

```{r}

# filter db, only observations where distance is > 0.2 mm

# gets new order 
imp_data$cycle <- factor(imp_data$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
imp_data$raw.material <- factor(imp_data$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))

counts.plot <-ggplot(data = imp_data) +  
              geom_point(mapping = aes(x = end, y = value, colour = raw.material), size = 0.5) +
              theme_classic() +
              labs(colour = "Raw material", x = "Distance (mm)", y = "Number of parts", title = "All raw materials") +
              facet_wrap(~ cycle)
              
print(counts.plot)

ggsave("../plots/counts.png")

# Flint
imp_data$cycle <- factor(imp_data$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
imp_data$raw.material <- factor(imp_data$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))

counts.plot.flt <-ggplot(subset(imp_data, raw.material=="flint")) +  
              geom_point(mapping = aes(x = end, y = value, colour = ID), size = 0.5) +
              theme_classic() +
              labs(colour = "ID", x = "Distance (mm)", y = "Number of parts", title = "Flint") +
              facet_wrap(~ cycle)
              
print(counts.plot.flt)

ggsave("../plots/flt.png")

# Quartzite
imp_data$cycle <- factor(imp_data$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
imp_data$raw.material <- factor(imp_data$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))


counts.plot.qrtz <-ggplot(subset(imp_data, raw.material=="quartzite")) +  
              geom_point(mapping = aes(x = end, y = value, colour = ID), size = 0.5) +
              theme_classic() +
              labs(colour = "ID", x = "Distance (mm)", y = "Number of parts", title =  "Quarzite") +
              facet_wrap(~ cycle)
              
print(counts.plot.qrtz)

ggsave("../plots/qrtz.png")

# Obsidian
imp_data$cycle <- factor(imp_data$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
imp_data$raw.material <- factor(imp_data$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))


counts.plot.obs <-ggplot(subset(imp_data, raw.material=="obsidian")) +  
              geom_point(mapping = aes(x = end, y = value, colour = ID), size = 0.5) +
              theme_classic() +
              labs(colour = "ID", x = "Distance (mm)", y = "Number of parts", title = "Obsidian") +
              facet_wrap(~ cycle)
              
print(counts.plot.obs)

ggsave("../plots/obs.png")

# Dacite
imp_data$cycle <- factor(imp_data$cycle, levels = c("0-125", "125-250", "250-500", "0-500"))
imp_data$raw.material <- factor(imp_data$raw.material, levels = c("flint", "obsidian", "dacite", "quartzite"))


counts.plot.dac <-ggplot(subset(imp_data, raw.material=="dacite")) +  
              geom_point(mapping = aes(x = end, y = value, colour = ID), size = 0.5) +
              theme_classic() +
              labs(colour = "ID", x = "Distance (mm)", y = "Number of parts", title = "Dacite") +
              facet_wrap(~ cycle)
              
print(counts.plot.dac)

ggsave("../plots/dac.png")

```

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

---

END OF SCRIPT

