---
title: "Import SMARTTESTER datasets"
author: "Joao Marreiros and David Nora"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```


---

# Goal of the script
This script imports and merges all single TXT-files (strokes + sensors) produced with the
Inotec Smarttester. The experiment involved 12 samples (3 samples from each 4 raw materials) which have been used in four cycles (0-250, 250-500, and 500-1000 strokes)
respectively. 
The script will:

1. Read in the original TXT-files   
2. Format and merge the data for each sample 
4. Combine the data from the 12 samples into one 
3. Write an XLSX-file and save an R object ready for further analysis in R 

This script is an adapted from...



```{r}
dir_in <- "../raw_data/"
dir_out <- "../derived_data/"

```


Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(tidyverse)
library(R.utils)
library(openxlsx)
library(tools)

```


---


# List all files and get names of the files  
```{r}
# List all CSV files in dir_in
TXT_files <- list.files(dir_in, pattern = "\\.txt$", recursive = TRUE, full.names = TRUE)

# Extract sample names from paths
samples_names <- dirname(dirname(dirname(TXT_files))) %>% # Path of folder 3 levels higher
                 basename() %>%                           # Name of folder 3 levels higher
                 unique()                                 # Unique names

```


# Define sensors 
```{r}
sensors <- data.frame(mess = paste0("Messung", 1:5), 
                      meas = c("Force", "Friction", "Depth", "Position", "Velocity"), 
                      unit = c("N", "N", "mm", "mm", "mm/s"))

```


# Merge all files and format the data 
```{r}
# Create named list, 1 element for each sample
sampl <- vector(mode = "list", length = length(samples_names)) 
names(sampl) <- samples_names

# For each sample
for (s in seq_along(samples_names)) {
  
  # Gets information through the path name and defines the cycle, raw material and 
  # contact material
  folder <- paste0(samples_names[s], "/") %>% 
            grep(TXT_files, value = TRUE) %>% 
            dirname() %>% 
            dirname() %>% 
            unique() %>% 
            basename() %>% 
            strsplit(., "_") 
  
  cycles <- sapply(folder, FUN = function(x) x[[3]])
  # Defines the number of the first stroke per cycle based on the name from the folders
  cycle_start <- gsub("-.*$", "", x = cycles) %>% 
                 # Converts into numeric             
                 as.numeric()
  
  # Orders the cycles
  order_cycles <- order(cycle_start)
  cycle_start <- cycle_start[order_cycles]
  cycle_start[1] <- 1
  cycles <- cycles[order_cycles]
  
  # Takes the information about the contact material
  cont_mat <- sapply(folder, FUN = function(x) x[[2]]) %>% 
              unique()
  
  # Takes the information about the raw material
  #raw_mat <- ifelse(grepl("FLT", names(sampl)[s]), "Flint", "Lydite")

  if(grepl("FLT", names(sampl)[s]) == TRUE) raw_mat <- "Flint"
  if(grepl("OBS", names(sampl)[s]) == TRUE) raw_mat <- "Obsidian"
  if(grepl("QTZ", names(sampl)[s]) == TRUE) raw_mat <- "Quatzite"
  if(grepl("DAC", names(sampl)[s]) == TRUE) raw_mat <- "Dacite"
  
  # Create named list, 1 element for each sensor ("Messung")
  sampl[[s]] <- vector(mode = "list", length = nrow(sensors))
  names(sampl[[s]]) <- sensors [["meas"]]
  
  # For each sensor ("Messung")
  for (m in seq_along(sampl[[s]])) {
      
    # Extract file names of all strokes for the given sensor
    # Paste sample name and slash to avoid partial matching
    s_m <- paste0(samples_names[[s]], "/") %>% 
           # Extract sample "s" from all files
           grep(TXT_files, value = TRUE) %>% 
           # Extract sensor "m" from sample "s"
           grep(sensors[["mess"]][m], ., value = TRUE) 
    
    # Create named list, 1 element for each stroke bin
    sampl[[s]][[m]] <- vector(mode = "list", length = length(cycles))
    names(sampl[[s]][[m]]) <- cycles
    
    # For each cycle
    for (cy in seq_along(sampl[[s]][[m]])) {
      
      # Extract file names of all strokes for each cycle
      s_m_cy <- grep(cycles[cy], s_m, value = TRUE)
     
      # Create named list, 1 element for each stroke
      sampl[[s]][[m]][[cy]] <- vector(mode = "list", length = length(s_m_cy))
      names(sampl[[s]][[m]][[cy]]) <- paste0("Stroke", seq_along(s_m_cy))
      
      # For each stroke
      for (st in seq_along(s_m_cy)) {
     
        # Read in TXT file
        sampl[[s]][[m]][[cy]][[st]] <- read.table(s_m_cy[st], skip = 4, sep = ";") %>% 
          
          # Add columns Step based on V2 and Stroke based on "st"
          mutate(Step = V2/100000+1, Stroke = st -1 + cycle_start[cy]) %>%    
          
          # Select columns stroke, step, V1
          select(Stroke, Step, V1)
        
        # Rename column V1 based on "m"
        names(sampl[[s]][[m]][[cy]][[st]])[3] <- sensors[m, "meas"] 
      }
      
      # rbind all files per cycle
      sampl[[s]][[m]][[cy]] <- do.call(rbind, sampl[[s]][[m]][[cy]])
    }
    
    # rbind all cycles per sensor
    sampl[[s]][[m]] <- do.call(rbind, sampl[[s]][[m]])
  }
  
  # rbind all sensors per sample
  sampl[[s]] <- full_join(sampl[[s]][[1]], sampl[[s]][[2]]) %>% 
    full_join(sampl[[s]][[3]]) %>% 
    full_join(sampl[[s]][[4]]) %>%
    full_join(sampl[[s]][[5]]) %>% 
    mutate(Sample = names(sampl)[s], Raw_material = raw_mat, 
           Contact_material = cont_mat) %>%
    
    select(Sample, Raw_material, Contact_material, everything())
}

# rbind all samples 
sampl <- do.call(rbind, sampl)

```



# Save data
## Format name of output file
```{r}
file_out <- "sampl"

```


## Write to XLSX
```{r}
write.xlsx(list(data = sampl, units = sensors), file = paste0(dir_out, file_out, ".xlsx"))

```


## Save R object
```{r}
saveObject(sampl, file = paste0(dir_out, file_out, ".Rbin"))

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```


---


END OF SCRIPT
