---
title: "Plots_inotec_data"
author: "Joao Marreiuros and David Nora"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---

# Goal of the script
This script plots all sensor data in order to visualizes the measurements 
recorded throughout the tool function experiment.
In this study the variable of interest is the *Penetration depth* 

```{r}
dir_in <- "../derived_data"
dir_out <- "../plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(patchwork)
library(doBy)
library(ggrepel)
library(openxlsx)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
info_in
```

# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

# replace "stroke" by "cycle"
colnames(imp_data)[colnames(imp_data) == "Stroke"] <- "Cycle"


```

The imported file is: "`r paste0("~/", data_file)`"  



# Plot each of the selected numeric variable 


## Plot showing the absolut penetration depths 
```{r}
# calculates the absolute depths reached per sample
abs.depth <- function(x) {
  noNA <- x[!is.na(x)]
  out <- abs(min(noNA) - max(noNA))
}

# Define grouping variable and compute the summary statistics 
depth <- summaryBy(Depth ~ Sample+Raw_material+Contact_material, 
                  data=imp_data, 
                  FUN=abs.depth)

str(depth)


depth[["Contact_material"]] <- factor(depth[["Contact_material"]])

# plots all depth points in one facet plot (contact material together)
p3 <- ggplot(data = depth, aes(x = Contact_material, 
                               y = Depth.abs.depth, colour = 
                                 Raw_material)) +
       geom_point() + labs(y = "Absolute depth (mm)") +
       facet_wrap(~Raw_material, strip.position = "bottom") +
       # avoids overplotting of the labels (sample IDs)
       geom_text_repel(aes(label=Sample), size = 2, 
                       nudge_x = -0.4, 
                       segment.size = 0.1, force = 2, 
                       seed = 123) +
       scale_y_continuous(trans = "reverse") +
       scale_x_discrete(position ="top") +
       # removes the "_" between "Contact_material in the legend 
       labs(x = "Contact material") + 
	     theme_classic() +
       theme(legend.position = "none") 
      
print(p3)

# save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), 
                   "_depth_a_plot_", ".pdf")
ggsave(filename = file_out, plot = p3, path = dir_out, 
       device = "pdf", 
       width = 25, height = 17, units = "cm")


depth[["Raw_material"]] <- factor(depth[["Raw_material"]])

# plots all depth points in one facet plot (contact material separated)
p4 <- ggplot(data = depth, aes(x = Contact_material, 
                               y = Depth.abs.depth, colour = 
                                 Raw_material)) +
       geom_point() + labs(y = "Absolute depth (mm)") +
       # avoids overplotting of the labels (sample IDs)
       geom_text_repel(aes(label=Sample), size = 2, 
                       nudge_x = -0.4, 
                       segment.size = 0.1, force = 2, 
                       seed = 123) +
       scale_y_continuous(trans = "reverse") +
       scale_x_discrete(position ="top") +
       # removes the "_" between "Contact_material in the legend 
       labs(x = "Contact material") + 
	     theme_classic() +
       theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
       theme(legend.position = "none") 
      
print(p4)

# save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), 
                   "_depth_b_plot_", ".pdf")
ggsave(filename = file_out, plot = p4, path = dir_out, 
       device = "pdf", 
       width = 25, height = 17, units = "cm")

```

## All sensor data
```{r}

sp <- split(imp_data, imp_data[["Sample"]])

for (i in seq_along(sp)) {
  # creates a sequence of every ~ 50th strokes 
  seq_st <- seq(1, length(unique(sp[[i]][["Cycle"]])), by = 40) %>% 
            c(max(unique(sp[[i]][["Cycle"]])))
  dat_i_all <- sp[[i]] %>% 
               filter(Cycle %in% seq_st)
  range_force_all <- range(dat_i_all[["Force"]])
  range_friction_all <- range(dat_i_all[["Friction"]])
  range_depth_all <- range(dat_i_all[["Depth"]])
  range_velocity_all <- range(dat_i_all[["Velocity"]])
       
  
    p1b <- ggplot(data = dat_i_all) +
        geom_line(aes(x = Step, y = Force, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Force [N]") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_force_all) +
        scale_x_continuous(breaks=c(1, 4, 7, 10, 15, 20, 25)) +
	      theme_classic()
  print(p1b)
  
    p2b <- ggplot(data = dat_i_all) +
        geom_line(aes(x = Step, y = Friction, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Friction [N]") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_friction_all) +
        scale_x_continuous(breaks=c(1, 4, 7, 10, 15, 20, 25)) +
	      theme_classic()
  print(p2b)
  
  p3b <- ggplot(data = dat_i_all) +
        geom_line(aes(x = Step, y = Depth, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Depth [mm]") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth_all) +
        scale_x_continuous(breaks=c(1, 4, 7, 10, 15, 20, 25)) +
	      theme_classic()
  print(p3b)
  
    p4b <- ggplot(data = dat_i_all) +
        geom_line(aes(x = Step, y = Velocity, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Velocity [mm/s]") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_velocity_all) +
        scale_x_continuous(breaks=c(1, 4, 7, 10, 15, 20, 25)) +
	      theme_classic()
  print(p4b)
  
  # patchwork plot
  pb <- p1b + p2b + p3b + p4b + plot_annotation(title = names(sp)[i]) + plot_layout(ncol = 1, guides = "collect")
  print(pb)
  # save to PDF
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_sensors_plot_", 
	                   names(sp)[i], ".pdf")
  ggsave(filename = file_out, plot = pb, path = dir_out, device = "pdf")
} 




```


## *Penetration depth* plots showing the strokes as lines 
```{r}
# plots all strokes per sample divided by 40 
# splits the data in the individual 24 samples
sp <- split(imp_data, imp_data[["Sample"]])


for (i in seq_along(sp)) {
  # creates a sequence of every ~ 50th cycles 
  seq_st <- seq(1, length(unique(sp[[i]][["Cycle"]])), by = 40) %>% 
            c(max(unique(sp[[i]][["Cycle"]])))
  dat_i_all <- sp[[i]] %>% 
               filter(Cycle %in% 1:500)
  range_depth <- range(dat_i_all[["Depth"]])
  p1 <- ggplot(data = dat_i_all, aes(x = Step, y = Depth, colour = Cycle)) +
        geom_line(aes(group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Depth (mm)") + ylab(NULL) +
        # reverses the legend starting with 0 going to 2000 strokes 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth) +
        # changes the 'Step-number' in the x-legend  
        theme_classic()
       
# plots only the first 125 cycles per sample  
  dat_i_250 <- sp[[i]] %>% 
              # takes only the first 50 cycles per sample
              filter(Cycle %in% 1:125)
  range_depth <- range(dat_i_all[["Depth"]])
  p2 <- ggplot(data = dat_i_250) +
        geom_line(aes(x = Step, y = Depth, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Depth (mm)") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth) +
	      theme_classic()

# plots only between 125 to 205 cycles per sample  
  dat_i_500 <- sp[[i]] %>% 
              # takes only the first 50 cycles per sample
              filter(Cycle %in% 126:250)
  range_depth <- range(dat_i_all[["Depth"]])
  p3 <- ggplot(data = dat_i_500) +
        geom_line(aes(x = Step, y = Depth, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Depth (mm)") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth) +
	      theme_classic()
  
  # plots only between 250 to 500 cycles per sample  
  dat_i_500 <- sp[[i]] %>% 
              # takes only the first 50 cycles per sample
              filter(Cycle %in% 251:500)
  range_depth <- range(dat_i_all[["Depth"]])
  p4 <- ggplot(data = dat_i_500) +
        geom_line(aes(x = Step, y = Depth, colour = Cycle, group = Cycle), alpha = 0.3) + 
        labs(x = "Step", y = "Depth (mm)") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth) +
	      theme_classic()  
  
  # patchwork plot
  p <- p2 + p3 + p4 + p1 +plot_annotation(title = names(sp)[i]) 
  print(p)

  # save to PDF
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_depth_plot_", 
	            names(sp)[i], ".pdf")
  ggsave(filename = file_out, plot = p, path = dir_out, 
         device = "pdf")
}

```


The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---
# Save data
## Write to XLSX (summary statistics)
```{r}
write.xlsx(list(depth = depth, depth_good = depth_good), 
                 file = paste0(dir_out, file_out, ".xlsx"))

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

---

END OF SCRIPT

