---
title: "Plots_aVSn-Inotec"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output: word_document
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```


# Goal of the script
This script plots  sensor data in order to visualizes the measurements 
recorded throughout the tool function experiment.  
Variables of interest are: 
* Penetration depth 



```{r}
dir_in <- "analysis/derived_data/aVSn/"
dir_out <- "analysis/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(patchwork)
library(doBy)
library(ggrepel)
library(openxlsx)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

```

The imported file is: "`r paste0("~/", data_file)`"  



# Plot each of the selected numeric variable 
## Plots showing the strokes as lines 
```{r}
# plots all 2000 strokes per sample divided by 40 
# splits the data in the individual 24 samples
sp <- split(imp_data, imp_data[["Sample"]])


for (i in seq_along(sp)) {
  # creates a sequence of every ~ 50th strokes 
  seq_st <- seq(1, length(unique(sp[[i]][["Stroke"]])), by = 40) %>% 
            c(max(unique(sp[[i]][["Stroke"]])))
  dat_i_all <- sp[[i]] %>% 
               filter(Stroke %in% seq_st)
  range_depth <- range(dat_i_all[["Depth"]])
  p1 <- ggplot(data = dat_i_all, aes(x = Step, y = Depth, colour = Stroke)) +
        geom_line(aes(group = Stroke), alpha = 0.3) + 
        labs(x = "Step", y = "Depth (mm)") + ylab(NULL) +
        # reverses the legend starting with 0 going to 2000 strokes 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth) +
        # changes the 'Step-number' in the x-legend  
        scale_x_continuous(breaks=c(1, 4, 7, 10)) +
	      theme_classic()
       
# plots only the first 50 strokes per sample  
  dat_i_50 <- sp[[i]] %>% 
              # takes only the first 50 strokes per sample
              filter(Stroke %in% 1:50)
  p2 <- ggplot(data = dat_i_50) +
        geom_line(aes(x = Step, y = Depth, colour = Stroke, group = Stroke), alpha = 0.3) + 
        labs(x = "Step", y = "Depth (mm)") + 
        scale_colour_continuous(trans = "reverse") + 
        coord_cartesian(ylim = range_depth) +
        scale_x_continuous(breaks=c(1, 4, 7, 10)) +
	      theme_classic()
  # patchwork plot
  p <- p2 + p1 + plot_annotation(title = names(sp)[i]) 
  print(p)

  # save to PDF
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_aVSn_plot_", 
	            names(sp)[i], ".pdf")
  ggsave(filename = file_out, plot = p, path = dir_out, 
         device = "pdf")
}

```



## Plot showing the absolut penetration depths 
### Plot of all samples 
```{r}
# calculates the absolute depths reached per sample
abs.depth <- function(x) {
  noNA <- x[!is.na(x)]
  out <- abs(min(noNA) - max(noNA))
}

# Define grouping variable and compute the summary statistics 
depth <- summaryBy(Depth ~ Sample+Raw_material+Contact_material, 
                  data=imp_data, 
                  FUN=abs.depth)

str(depth)


depth[["Contact_material"]] <- factor(depth[["Contact_material"]])
# GrandBudapest1 = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236")
custom.col3 <- data.frame(type = levels(depth$Contact_material), 
                         col = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236")) 
depth$col <- custom.col3[depth$Contact_material, "col"]


# plots all depth points in one facet plot (contact material together)
p3 <- ggplot(data = depth, aes(x = Contact_material, 
                               y = Depth.abs.depth, colour = 
                                 Contact_material)) +
       geom_point() + labs(y = "Absolute depth (mm)") +
       scale_colour_manual(values = custom.col3$col) + 
       facet_wrap(~Raw_material, strip.position = "bottom") +
       # avoids overplotting of the labels (sample IDs)
       geom_text_repel(aes(label=Sample), size = 2, 
                       nudge_x = -0.4, 
                       segment.size = 0.1, force = 2, 
                       seed = 123) +
       scale_y_continuous(trans = "reverse") +
       scale_x_discrete(position ="top") +
       # removes the "_" between "Contact_material in the legend 
       labs(x = "Contact material") + 
	     theme_classic() +
       theme(legend.position = "none") 
      
print(p3)

# save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), 
                   "_aVSn_P3_depth_plot_", ".pdf")
ggsave(filename = file_out, plot = p3, path = dir_out, 
       device = "pdf", 
       width = 25, height = 17, units = "cm")


depth[["Raw_material"]] <- factor(depth[["Raw_material"]])
#Royal1 = c("#899DA4", "#C93312", "#FAEFD1", "#DC863B")
custom.col7 <- data.frame(type = levels(depth$Raw_material), 
                         col = c("#899DA4", "#DC863B")) 
depth$col <- custom.col7[depth$Raw_material, "col"]


# plots all depth points in one facet plot (contact material separated)
p4 <- ggplot(data = depth, aes(x = Contact_material, 
                               y = Depth.abs.depth, colour = 
                                 Raw_material)) +
       geom_point() + labs(y = "Absolute depth (mm)") +
       scale_colour_manual(values = custom.col7$col) + 
       facet_wrap(~Contact_material, strip.position = "bottom") +
       # avoids overplotting of the labels (sample IDs)
       geom_text_repel(aes(label=Sample), size = 2, 
                       nudge_x = -0.4, 
                       segment.size = 0.1, force = 2, 
                       seed = 123) +
       scale_y_continuous(trans = "reverse") +
       scale_x_discrete(position ="top") +
       # removes the "_" between "Contact_material in the legend 
       labs(x = "Contact material") + 
	     theme_classic() +
       theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
       theme(legend.position = "none") 
      
print(p4)

# save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), 
                   "_aVSn_P4_depth_plot_", ".pdf")
ggsave(filename = file_out, plot = p4, path = dir_out, 
       device = "pdf", 
       width = 25, height = 17, units = "cm")

```


The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---
# Save data
## Write to XLSX (summary statistics)
```{r}
write.xlsx(list(depth = depth, depth_good = depth_good), 
                 file = paste0(dir_out, file_out, ".xlsx"))

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT

