---
title: "Equotip hardness measurements"
author: "Joao Marreiros and David Nora"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_knit$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```

---


# Goal of the script
This script reads the csv file (measurements have been generated with the Equotip Leeb C rebound)  and formats the data for a statistical analysis.  
The script will:

1. Read in the original csv file and organise the data 
2. Plot the data 
3. Write an XLSX-file and save an R object ready for further analysis in R 


Imported files are in: ‘../analysis/raw_data’

Figures are saved in: ‘../analysis/plots’

Tables are saved in: ‘../analysis/derived_data’

---


# Load packages
```{r}
library(tidyverse)
library(AICcmodavg)
library(ggplot2)
library(utils)
library(qwraps2)
library(kableExtra)
library(doBy)

```

---

# Read in original xlsx-file
```{r}
# List all CSV files in dir_in

imp_data <- read_csv("../raw_data/data.csv")

```

# Organize data
```{r}
# organizing data
longdata <- imp_data %>%
  gather("M1", "M2","M3", "M4", "M5","M6","M7", "M8", "M9", "M10", key = Measurment, value = HLC)

longdata

write_csv(longdata, "../derived_data/longdata.csv")

```

# Data analsysis - descriptive stats  
```{r}
# descriptive statistics 

nminmaxmeanmedsd <- function(x){
	y <- x[!is.na(x)]
	n_test <- length(y)
	min_test <- min(y)
	max_test <- max(y)
	mean_test <- mean(y)
 	med_test <- median(y)
 	sd_test <- sd(y)
 	out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test)
 	names(out) <- c("n", "min", "max", "mean", "median", "sd")
 	return(out)
}

num.var <- 6:length(longdata)
stats <- summaryBy(.~rawmaterial, data=longdata[c("rawmaterial", names(longdata)[num.var])], FUN=nminmaxmeanmedsd)

stats

write_csv(stats, "../derived_data/stats.csv")



```

# ANOVA analysis
```{r}
# anova
#longdata$rawmaterial <- as.factor(longdata$rawmaterial)

anova <- aov(HLC ~ rawmaterial, data = longdata)
anovafactor <- summary(aov(HLC ~ factor(rawmaterial), data = longdata))
one <- oneway.test(HLC ~ rawmaterial, data = longdata)
tuk <- TukeyHSD(aov(HLC ~ factor(rawmaterial), data = longdata))

anova
anovafactor
one
tuk

```


# Data analsysis - plot 
```{r}
# boxplot
data_plot <- ggplot (longdata, aes(rawmaterial, HLC, color = rawmaterial)) + 
             theme_classic() + 
             theme(legend.title = element_blank()) +
             geom_boxplot() +
             geom_jitter() + labs(x="Raw material", y="Leeb Rebound Hardness in HLC", title="") 

print(data_plot)

ggsave("../plots/dataplot.png")


```


# sessionInfo() and RStudio version
```{r}
sessionInfo()

```
